---
title: "DS4SI-final"
author: "Group H"
date: "12/05/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = T,
                      message = F,
                      warning = F,
                      results = "asis",
                      cache = T,
                      cache.lazy = FALSE)

library(weights)
library(tidyverse)
library(foreign)
library(ggrepel)
library(lme4)

```

```{r data preparation}
# the data files are too large to be put in the repository, specify your own data repository
# data_file_path <- "/Users/michaelfive/Google Drive/PISA/Task Flow/replication15feb"
# 
# ## load in PISA 2015 data
# dat <- read.spss(file.path(data_file_path, "CY6_MS_CMB_STU_QQQ.sav"), to.data.frame=T)
# 
# # load in GGI 2015 data
# cdat <- read.csv(file.path(data_file_path,"GGI2015.csv"), as.is = T)

load("~/Desktop/R Directory/Git learning/DS4SI-final/DS4SI-final.RData")
```

```{r functions}
# functions to calculate cohen's d
cohens_d <- function(x, y) {
  lx <- length(x)- 1
  ly <- length(y)- 1
  md  <- mean(x, na.rm = T) - mean(y, na.rm = T)        ## mean difference (numerator)
  csd <- lx * var(x,na.rm = T) + ly * var(y,na.rm = T)
  csd <- csd/(lx + ly)
  csd <- sqrt(csd)                     ## common sd computation
  
  cd  <- md/csd                        ## cohen's d
  return(cd)
}

```

# Replication: PISA & GGGI

We replicate here all the analysis in the original study that used the PISA 2015 data and the GGGI 2015 data.

```{r Fig 3 Pabel A}

## select relevant variables and countries from the data set

dat.s <- dat %>% select(CNTRYID,W_FSTUWT,ST004D01T, SCIEEFF,PV1MATH:PV10SCIE) 
names(dat.s)[c(1:4)] <- c("country","weight","gender", "efficacy")

# turn factor into numeric
dat.s$efficacy <- as.numeric(as.character(dat.s$efficacy))

dat.s <- dat.s[!dat.s$country %in% c( 
  "Chinese Taipei","Puerto Rico",
  "Dominican Republic", "Kosovo",
  "USA (Massachusetts)", "USA (North Carolina)",
  "Argentina (Ciudad Autónoma de Buenos)",
  "Spain (Regions)","Spain (Regions)_duplicated_971") ,  ]

## calculation of the intraindividual scores

cat("Step 1: We standardized the mathematics, science, and reading scores on a nation-by-nation basis. We call these new standardized scores zMath, zRead- ing, and zScience, respectively.")

dat.s$scie <- rowMeans(dat.s[,paste("PV",1:5,"SCIE",sep="")])
dat.s$math <- rowMeans(dat.s[,paste("PV",1:5,"MATH",sep="")])
dat.s$read <- rowMeans(dat.s[,paste("PV",1:5,"READ",sep="")])

dat.s <- dat.s %>% group_by(country) %>% mutate(
  z_scie = stdz(scie, weight),
  z_math = stdz(math, weight),
  z_read = stdz(read, weight)
)

cat("Step 2: We standardized the mathematics, science, and reading scores on a nation-by-nation basis. We call these new standardized scores zMath, zReading, and zScience, respectively.")

dat.s$general <- rowMeans( dat.s[,c("z_math","z_read","z_scie")] ) 

dat.s <- dat.s %>% group_by(country) %>% mutate( gz = stdz(general, weight) )

cat("Step 3: Then, we calculated each student’s intraindivid- ual strengths by subtracting zGeneral as follows: relative science strength = zScience – zGeneral, relative math strength = zMath – zGeneral, rela- tive reading strength = zReading – zGeneral.")

dat.s$iiScie <- dat.s$z_scie - dat.s$gz
dat.s$iiMath <- dat.s$z_math - dat.s$gz
dat.s$iiRead <- dat.s$z_read - dat.s$gz

dat.s <- dat.s %>% group_by(country) %>% mutate(
  iiSciez = stdz(iiScie, weight),
  iiMathz = stdz(iiMath, weight),
  iiReadz = stdz(iiRead, weight)
)

cat("Step 4: Then, we calculated each student’s intraindividual strengths by subtracting zGeneral as follows: relative science strength = zScience – zGeneral, relative math strength = zMath – zGeneral, rela- tive reading strength = zReading – zGeneral.")

dat.s <- select(dat.s, country, weight, gender, scie, math, read, efficacy, iiSciez, iiMathz, iiReadz , gz )

cdat.s <- dat.s %>% group_by(country) %>% dplyr::summarize(
     scieGap = weighted.mean(iiSciez[gender == "Male"], w = weight[gender == "Male"]) - weighted.mean(iiSciez[gender == "Female"], w = weight[gender == "Female"]),
     mathGap = weighted.mean(iiMathz[gender == "Male"], w = weight[gender == "Male"]) - weighted.mean(iiMathz[gender == "Female"], w = weight[gender == "Female"]),
     readGap = weighted.mean(iiReadz[gender == "Male"], w = weight[gender == "Male"]) - weighted.mean(iiReadz[gender == "Female"], w = weight[gender == "Female"])
)

## now we load in the GGI data and combine with gap pisa results
cdat.n <- left_join(cdat.s, cdat , by = "country")

## make data frame for easier use with cor.test
tmp = as.data.frame( cdat.n )
tmp = tmp[ !is.na(tmp[,"GGI"]),]

## report the N so you can compare with the paper (this should be 62)
cat(paste("Countries with both pisa data and gender equality data:",sum(!is.na(tmp[,"scieGap"])&!is.na(tmp[,"GGI"]))))

## print the correlation of Figure 3A and page 585
print ( cor.test( tmp[,"scieGap"] , tmp[,"GGI"] , method="spearman") )

## print the reading correlation, as reported on page 585
print ( cor.test( tmp[,"readGap"] , tmp[,"GGI"] , method="spearman") )


## plot
ggplot(tmp, aes(y = GGI, x = scieGap)) + 
  geom_point(color = "red", size = 0.8) + 
  geom_smooth(method = "lm", fill = NA, size = 0.8) + 
  geom_text_repel(aes(label = country)) +
  labs(y = "Global Gender Gap Index (GGGI)", x = "Gender Gap in Intraindividual Science Performance (d)") +
  theme_classic()

```

```{r Fig 2 Panel A&B, fig.height = 10, fig.width = 8}
## Panel A
cdat.n.cohend <- dat.s %>% group_by(country) %>% 
  dplyr::summarize(
    
    scieGap = cohens_d(scie[gender == "Male"], scie[gender == "Female"]),
    mathGap = cohens_d(math[gender == "Male"], math[gender == "Female"]),
    readGap = cohens_d(read[gender == "Male"], read[gender == "Female"]),
    efficacyGap = cohens_d(efficacy[gender == "Male"], efficacy[gender == "Female"])
    
  ) %>%
  left_join(., cdat , by = "country")

cdat.n.cohend.long <- cdat.n.cohend %>% 
  na.omit() %>%
  gather(scieGap:readGap,key = "gap_type", value = "gap")


ggplot(cdat.n.cohend.long, aes(x = reorder(country, desc(country)), y = gap, fill = gap_type)) +
  geom_bar(stat = "identity", position = "dodge") + 
  scale_fill_manual(values=c("#00BA38", "#F8766D", "#619CFF")) +
  coord_flip() +
  labs(x = "", y = "Sex differences in National PISA Scores (d)",
       fill = "") + 
  theme_classic()


## Panel B
cdat.n.long <- gather(na.omit(cdat.n), scieGap:readGap,key = "gap_type", value = "gap")

ggplot(cdat.n.long, aes(x = reorder(country, desc(country)), y = gap, fill = gap_type)) +
  geom_bar(stat = "identity", position = "dodge") + 
  scale_fill_manual(values=c("#00BA38", "#F8766D", "#619CFF")) +
  coord_flip() +
  labs(x = "", y = "Sex differences in Intraindividual Scores (d)",
       fill = "") + 
  theme_classic()
```

```{r Fig 4}
ggplot(cdat.n.cohend, aes(x = GGI, y = efficacyGap)) +
  geom_point(color = "red", size = 0.8) + 
  geom_smooth(method = "lm", fill = NA, size = 0.8) + 
  geom_text_repel(aes(label = country)) +
  labs(x = "Global Gender Gap Index (GGGI)", y = "Sex Differences in Science Self-Efficacy (d)") +
  theme_classic()

```


# Additional Analysis 

First, we try to predict GGGI at the country level with science relative performance gap, controlling for the other gaps in math and reading. Results show that science relative performance gap indeed predicts GGGI even after controlling for the other gaps. This may indicate that the explanation of the "gender gap paradox" could really be reversed as no causal path is evident from correlational models.

```{r analysis 1}

lm1 <- lm(GGI ~ scieGap, cdat.n)
lm2 <- lm(GGI ~ scieGap + mathGap, cdat.n)
lm3 <- lm(GGI ~ scieGap + mathGap + readGap, cdat.n)

texreg::htmlreg(list(lm1, lm2, lm3))
```


Second, we try using the plausible values correctly (i.e. averaging the correlation estimates rather than the plausible values before further analysis) and repeat their z-scoring procedure. We arrive at `r` = 0.405, which is pretty close to the original results.

```{r analysis 2}
## select relevant variables and countries from the data set

dat.s <- dat %>% select(CNTRYID,W_FSTUWT,ST004D01T, SCIEEFF,PV1MATH:PV10SCIE) 
names(dat.s)[c(1:4)] <- c("country","weight","gender", "efficacy")

# turn factor into numeric
dat.s$efficacy <- as.numeric(as.character(dat.s$efficacy))

dat.s <- dat.s[!dat.s$country %in% c( 
  "Chinese Taipei","Puerto Rico",
  "Dominican Republic", "Kosovo",
  "USA (Massachusetts)", "USA (North Carolina)",
  "Argentina (Ciudad Autónoma de Buenos)",
  "Spain (Regions)","Spain (Regions)_duplicated_971") ,  ]

get.cor1 <- function(dat.s, i){

    dat.s$scie <- dat.s[,paste("PV",i,"SCIE",sep="")]
    dat.s$math <- dat.s[,paste("PV",i,"MATH",sep="")]
    dat.s$read <- dat.s[,paste("PV",i,"READ",sep="")]
    
    dat.s <- dat.s %>% group_by(country) %>% mutate(
      z_scie = stdz(scie, weight),
      z_math = stdz(math, weight),
      z_read = stdz(read, weight)
    )
    
    dat.s$general <- rowMeans( dat.s[,c("z_math","z_read","z_scie")] ) 
    
    dat.s <- dat.s %>% group_by(country) %>% mutate( gz = stdz(general, weight) )
    
    dat.s$iiScie <- dat.s$z_scie - dat.s$gz
    dat.s$iiMath <- dat.s$z_math - dat.s$gz
    dat.s$iiRead <- dat.s$z_read - dat.s$gz
    
    dat.s <- dat.s %>% group_by(country) %>% mutate(
      iiSciez = stdz(iiScie, weight),
      iiMathz = stdz(iiMath, weight),
      iiReadz = stdz(iiRead, weight)
    )
    
    dat.s <- select(dat.s, country, weight, gender, scie, math, read, efficacy, iiSciez, iiMathz, iiReadz , gz )
    
    cdat.s <- dat.s %>% group_by(country) %>% dplyr::summarize(
         scieGap = weighted.mean(iiSciez[gender == "Male"], w = weight[gender == "Male"]) - weighted.mean(iiSciez[gender == "Female"], w = weight[gender == "Female"]),
         mathGap = weighted.mean(iiMathz[gender == "Male"], w = weight[gender == "Male"]) - weighted.mean(iiMathz[gender == "Female"], w = weight[gender == "Female"]),
         readGap = weighted.mean(iiReadz[gender == "Male"], w = weight[gender == "Male"]) - weighted.mean(iiReadz[gender == "Female"], w = weight[gender == "Female"])
    )
    
    ## now we load in the GGI data and combine with gap pisa results
    cdat.n <- left_join(cdat.s, cdat , by = "country")
    
    ## make data frame for easier use with cor.test
    tmp = as.data.frame( cdat.n )
    tmp = tmp[ !is.na(tmp[,"GGI"]),]
    
    ## print the correlation of Figure 3A and page 585
    cor <- suppressWarnings(cor.test( tmp[,"scieGap"] , tmp[,"GGI"] , method="spearman"))
    
    ## return correlation coefficient
    return(cor$estimate)

}

# run a for loop to get all 10 estimates
df <- c()
for (i in 1:10) {
  
  df[i] <- get.cor1(dat.s, i)
  
}

# average the estimates
mean(df) # 0.4045611
```

Third, we try a different way of calculating relative strength. We first calculate the relative strength by subtracting science score from the mean score of science, reading and math for each individual. We then standardize this score within each country. We arrive at `r` = 0.401, which is also pretty close to the original results.


```{r analysis 3}

get.cor2 <- function(dat.s, i){

    dat.s$scie <- dat.s[,paste("PV",i,"SCIE",sep="")]
    dat.s$math <- dat.s[,paste("PV",i,"MATH",sep="")]
    dat.s$read <- dat.s[,paste("PV",i,"READ",sep="")]
    
    dat.s$general <- rowMeans( dat.s[,c("math","read","scie")] ) 
    
    dat.s$iiScie <- dat.s$scie - dat.s$general
    dat.s$iiMath <- dat.s$math - dat.s$general
    dat.s$iiRead <- dat.s$read - dat.s$general
    
    dat.s <- dat.s %>% group_by(country) %>% mutate(
      iiSciez = stdz(iiScie, weight),
      iiMathz = stdz(iiMath, weight),
      iiReadz = stdz(iiRead, weight)
    )
    
    dat.s <- select(dat.s, country, weight, gender, scie, math, read, efficacy, iiSciez, iiMathz, iiReadz)
    
    cdat.s <- dat.s %>% group_by(country) %>% dplyr::summarize(
         scieGap = weighted.mean(iiSciez[gender == "Male"], w = weight[gender == "Male"]) - weighted.mean(iiSciez[gender == "Female"], w = weight[gender == "Female"]),
         mathGap = weighted.mean(iiMathz[gender == "Male"], w = weight[gender == "Male"]) - weighted.mean(iiMathz[gender == "Female"], w = weight[gender == "Female"]),
         readGap = weighted.mean(iiReadz[gender == "Male"], w = weight[gender == "Male"]) - weighted.mean(iiReadz[gender == "Female"], w = weight[gender == "Female"])
    )
    
    ## now we load in the GGI data and combine with gap pisa results
    cdat.n <- left_join(cdat.s, cdat , by = "country")
    
    ## make data frame for easier use with cor.test
    tmp = as.data.frame( cdat.n )
    tmp = tmp[ !is.na(tmp[,"GGI"]),]
    
    ## print the correlation of Figure 3A and page 585
    cor <- suppressWarnings(cor.test( tmp[,"scieGap"] , tmp[,"GGI"] , method="spearman"))
    
    ## return correlation coefficient
    return(cor$estimate)

}

# run a for loop to get all 10 estimates
df <- c()
for (i in 1:10) {
  
  df[i] <- get.cor2(dat.s, i)
  
}

# average the estimates
mean(df) # 0.4014356
```

Forth and Fifth, we build multilevel models to predict (1) raw individual perforamnce score and (2) the relative strength score using country-level GGGI, controlling for other individual level confounders. Using only the first plausible score (a recommended unbiased shortcut as recommended by the PISA data manual), We found that GGGI does not interact with gender to predict either the raw science performance score or the science performance relative to the overall performance in science, reading and math. This finding provides is contrary to the main findings in the original paper, potentially because we considered the multilevel structure of the data. Importantly, although we attempted to add individual level covariates, the models mostly failed to converge.

```{r analysis 4&5}
dat.s <- dat %>% select(CNTRYID, CNTSCHID, CNTSTUID,
                            ST001D01T,ST004D01T, SMINS, 
                            TMINS, ESCS, REPEAT,
                            IMMIG, JOYSCIE, INTBRSCI, 
                            INSTSCIE,SCIEEFF,EPIST, 
                            SCIEACT, MOTIVAT, W_FSTUWT, PV1SCIE, PV1READ, PV1MATH) 

names(dat.s) <- c("country", "schid", "id",
                "grade","gender", "science_time",
                "total_time","ses", "graderepeat", 
                "immigration", "sci_enjoy","sci_interest", 
                "sci_instrument","sci_efficacy","sci_epist",
                "sci_activity", "sci_motivation", "weight", "pv1scie", "pv1read", "pv1math"
                )

dat.s$scie_rel <- rowMeans(dat.s[,c("pv1scie", "pv1read", "pv1math")], na.rm = T)
dat.s$scie_rel <- dat.s$pv1scie - dat.s$scie_rel


dat.s <- left_join(dat.s, cdat, by = "country")

fit1 <- lmer(pv1scie ~ 1 + gender + 
              (1 | schid:country) + # level 2 random effect
              (1 + gender | country ), # level 3 random effect
           weights = weight, data = dat.s)

fit2 <- lmer(pv1scie ~ 1 + gender * GGI + 
              (1 | schid:country) + # level 2 random effect
              (1 + gender | country ), # level 3 random effect
           weights = weight, data = dat.s)

fit3 <- lmer(scie_rel ~ 1 + gender + 
              (1 | schid:country) + # level 2 random effect
              (1 + gender | country ), # level 3 random effect
           weights = weight, data = dat.s)

fit4 <- lmer(scie_rel ~ 1 + gender * GGI + 
              (1 | schid:country) + # level 2 random effect
              (1 + gender | country ), # level 3 random effect
           weights = weight, data = dat.s)

# model won't converge well with covariates


texreg::htmlreg(list(fit1, fit2, fit3, fit4),
                  custom.model.names = c("Raw score", "Raw score",
                                         "Relative score", "Relative score"),
                  digits = 3)


```
