---
title: "DS4SI-final"
author: "Michael Wu"
date: "11/30/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(weights)
library(tidyverse)
library(foreign)
library(ggrepel)
library(lme4)

```

```{r data preparation}
# the data files are too large to be put in the repository, specify your own data repository
data_file_path <- "/Users/michaelfive/Google Drive/PISA/Task Flow/replication15feb"

## load in PISA 2015 data
dat <- read.spss(file.path(data_file_path, "CY6_MS_CMB_STU_QQQ.sav"), to.data.frame=T)

# load in GGI 2015 data
cdat <- read.csv(file.path(data_file_path,"GGI2015.csv"), as.is = T)

```

```{r functions}
# functions for cohen's d
cohens_d <- function(x, y) {
  lx <- length(x)- 1
  ly <- length(y)- 1
  md  <- mean(x, na.rm = T) - mean(y, na.rm = T)        ## mean difference (numerator)
  csd <- lx * var(x,na.rm = T) + ly * var(y,na.rm = T)
  csd <- csd/(lx + ly)
  csd <- sqrt(csd)                     ## common sd computation
  
  cd  <- md/csd                        ## cohen's d
  return(cd)
}

```

# Replication: PISA & GGGI

We replicate here all the analysis in the original study that used the PISA 2015 data and the GGGI 2015 data.

```{r Fig 3 Pabel A}

## select relevant variables and countries from the data set

dat.s <- dat %>% select(CNTRYID,W_FSTUWT,ST004D01T, SCIEEFF,PV1MATH:PV10SCIE) 
names(dat.s)[c(1:4)] <- c("country","weight","gender", "efficacy")

# turn factor into numeric
dat.s$efficacy <- as.numeric(as.character(dat.s$efficacy))

dat.s <- dat.s[!dat.s$country %in% c( 
  "Chinese Taipei","Puerto Rico",
  "Dominican Republic", "Kosovo",
  "USA (Massachusetts)", "USA (North Carolina)",
  "Argentina (Ciudad Autónoma de Buenos)",
  "Spain (Regions)","Spain (Regions)_duplicated_971") ,  ]

## calculation of the intraindividual scores

cat("Step 1: We standardized the mathematics, science, and reading scores on a nation-by-nation basis. We call these new standardized scores zMath, zRead- ing, and zScience, respectively.")

dat.s$scie <- rowMeans(dat.s[,paste("PV",1:5,"SCIE",sep="")])
dat.s$math <- rowMeans(dat.s[,paste("PV",1:5,"MATH",sep="")])
dat.s$read <- rowMeans(dat.s[,paste("PV",1:5,"READ",sep="")])

dat.s <- dat.s %>% group_by(country) %>% mutate(
  z_scie = stdz(scie, weight),
  z_math = stdz(math, weight),
  z_read = stdz(read, weight)
)

cat("Step 2: We standardized the mathematics, science, and reading scores on a nation-by-nation basis. We call these new standardized scores zMath, zReading, and zScience, respectively.")

dat.s$general <- rowMeans( dat.s[,c("z_math","z_read","z_scie")] ) 

dat.s <- dat.s %>% group_by(country) %>% mutate( gz = stdz(general, weight) )

cat("Step 3: Then, we calculated each student’s intraindivid- ual strengths by subtracting zGeneral as follows: relative science strength = zScience – zGeneral, relative math strength = zMath – zGeneral, rela- tive reading strength = zReading – zGeneral.")

dat.s$iiScie <- dat.s$z_scie - dat.s$gz
dat.s$iiMath <- dat.s$z_math - dat.s$gz
dat.s$iiRead <- dat.s$z_read - dat.s$gz

dat.s <- dat.s %>% group_by(country) %>% mutate(
  iiSciez = stdz(iiScie, weight),
  iiMathz = stdz(iiMath, weight),
  iiReadz = stdz(iiRead, weight)
)

cat("Step 4: Then, we calculated each student’s intraindividual strengths by subtracting zGeneral as follows: relative science strength = zScience – zGeneral, relative math strength = zMath – zGeneral, rela- tive reading strength = zReading – zGeneral.")

dat.s <- select(dat.s, country, weight, gender, scie, math, read, efficacy, iiSciez, iiMathz, iiReadz , gz )

cdat.s <- dat.s %>% group_by(country) %>% dplyr::summarize(
     scieGap = weighted.mean(iiSciez[gender == "Male"], w = weight[gender == "Male"]) - weighted.mean(iiSciez[gender == "Female"], w = weight[gender == "Female"]),
     mathGap = weighted.mean(iiMathz[gender == "Male"], w = weight[gender == "Male"]) - weighted.mean(iiMathz[gender == "Female"], w = weight[gender == "Female"]),
     readGap = weighted.mean(iiReadz[gender == "Male"], w = weight[gender == "Male"]) - weighted.mean(iiReadz[gender == "Female"], w = weight[gender == "Female"])
)

## now we load in the GGI data and combine with gap pisa results
cdat.n <- left_join(cdat.s, cdat , by = "country")

## make data frame for easier use with cor.test
tmp = as.data.frame( cdat.n )
tmp = tmp[ !is.na(tmp[,"GGI"]),]

## report the N so you can compare with the paper (this should be 62)
cat(paste("Countries with both pisa data and gender equality data:",sum(!is.na(tmp[,"scieGap"])&!is.na(tmp[,"GGI"]))))

## print the correlation of Figure 3A and page 585
print ( cor.test( tmp[,"scieGap"] , tmp[,"GGI"] , method="spearman") )

## print the reading correlation, as reported on page 585
print ( cor.test( tmp[,"readGap"] , tmp[,"GGI"] , method="spearman") )


## plot
ggplot(tmp, aes(y = GGI, x = scieGap)) + 
  geom_point(color = "red", size = 0.8) + 
  geom_smooth(method = "lm", fill = NA, size = 0.8) + 
  geom_text_repel(aes(label = country)) +
  labs(y = "Global Gender Gap Index (GGGI)", x = "Gender Gap in Intraindividual Science Performance (d)") +
  theme_classic()

```

```{r Fig 2 Panel A&B}
## Panel A
cdat.n.cohend <- dat.s %>% group_by(country) %>% 
  dplyr::summarize(
    
    scieGap = cohens_d(scie[gender == "Male"], scie[gender == "Female"]),
    mathGap = cohens_d(math[gender == "Male"], math[gender == "Female"]),
    readGap = cohens_d(read[gender == "Male"], read[gender == "Female"]),
    efficacyGap = cohens_d(efficacy[gender == "Male"], efficacy[gender == "Female"])
    
  ) %>%
  left_join(., cdat , by = "country")

cdat.n.cohend.long <- cdat.n.cohend %>% 
  na.omit() %>%
  gather(scieGap:readGap,key = "gap_type", value = "gap")


ggplot(cdat.n.cohend.long, aes(x = reorder(country, desc(country)), y = gap, fill = gap_type)) +
  geom_bar(stat = "identity", position = "dodge") + 
  scale_fill_manual(values=c("#00BA38", "#F8766D", "#619CFF")) +
  coord_flip() +
  labs(x = "", y = "Sex differences in National PISA Scores (d)",
       fill = "") + 
  theme_classic()


## Panel B
cdat.n.long <- gather(na.omit(cdat.n), scieGap:readGap,key = "gap_type", value = "gap")

ggplot(cdat.n.long, aes(x = reorder(country, desc(country)), y = gap, fill = gap_type)) +
  geom_bar(stat = "identity", position = "dodge") + 
  scale_fill_manual(values=c("#00BA38", "#F8766D", "#619CFF")) +
  coord_flip() +
  labs(x = "", y = "Sex differences in Intraindividual Scores (d)",
       fill = "") + 
  theme_classic()
```

```{r Fig 4}
ggplot(cdat.n.cohend, aes(x = GGI, y = efficacyGap)) +
  geom_point(color = "red", size = 0.8) + 
  geom_smooth(method = "lm", fill = NA, size = 0.8) + 
  geom_text_repel(aes(label = country)) +
  labs(x = "Global Gender Gap Index (GGGI)", y = "Sex Differences in Science Self-Efficacy (d)") +
  theme_classic()

```


# Additional Analysis 

First, we try using the plausible values correctly (i.e. averaging the correlation estimates rather than the plausible values before further analysis) and repeat their z-scoring procedure.

```{r analysis 1}
## select relevant variables and countries from the data set

dat.s <- dat %>% select(CNTRYID,W_FSTUWT,ST004D01T, SCIEEFF,PV1MATH:PV10SCIE) 
names(dat.s)[c(1:4)] <- c("country","weight","gender", "efficacy")

# turn factor into numeric
dat.s$efficacy <- as.numeric(as.character(dat.s$efficacy))

dat.s <- dat.s[!dat.s$country %in% c( 
  "Chinese Taipei","Puerto Rico",
  "Dominican Republic", "Kosovo",
  "USA (Massachusetts)", "USA (North Carolina)",
  "Argentina (Ciudad Autónoma de Buenos)",
  "Spain (Regions)","Spain (Regions)_duplicated_971") ,  ]

get.cor1 <- function(dat.s, i){

    dat.s$scie <- dat.s[,paste("PV",i,"SCIE",sep="")]
    dat.s$math <- dat.s[,paste("PV",i,"MATH",sep="")]
    dat.s$read <- dat.s[,paste("PV",i,"READ",sep="")]
    
    dat.s <- dat.s %>% group_by(country) %>% mutate(
      z_scie = stdz(scie, weight),
      z_math = stdz(math, weight),
      z_read = stdz(read, weight)
    )
    
    dat.s$general <- rowMeans( dat.s[,c("z_math","z_read","z_scie")] ) 
    
    dat.s <- dat.s %>% group_by(country) %>% mutate( gz = stdz(general, weight) )
    
    dat.s$iiScie <- dat.s$z_scie - dat.s$gz
    dat.s$iiMath <- dat.s$z_math - dat.s$gz
    dat.s$iiRead <- dat.s$z_read - dat.s$gz
    
    dat.s <- dat.s %>% group_by(country) %>% mutate(
      iiSciez = stdz(iiScie, weight),
      iiMathz = stdz(iiMath, weight),
      iiReadz = stdz(iiRead, weight)
    )
    
    dat.s <- select(dat.s, country, weight, gender, scie, math, read, efficacy, iiSciez, iiMathz, iiReadz , gz )
    
    cdat.s <- dat.s %>% group_by(country) %>% dplyr::summarize(
         scieGap = weighted.mean(iiSciez[gender == "Male"], w = weight[gender == "Male"]) - weighted.mean(iiSciez[gender == "Female"], w = weight[gender == "Female"]),
         mathGap = weighted.mean(iiMathz[gender == "Male"], w = weight[gender == "Male"]) - weighted.mean(iiMathz[gender == "Female"], w = weight[gender == "Female"]),
         readGap = weighted.mean(iiReadz[gender == "Male"], w = weight[gender == "Male"]) - weighted.mean(iiReadz[gender == "Female"], w = weight[gender == "Female"])
    )
    
    ## now we load in the GGI data and combine with gap pisa results
    cdat.n <- left_join(cdat.s, cdat , by = "country")
    
    ## make data frame for easier use with cor.test
    tmp = as.data.frame( cdat.n )
    tmp = tmp[ !is.na(tmp[,"GGI"]),]
    
    ## print the correlation of Figure 3A and page 585
    cor <- suppressWarnings(cor.test( tmp[,"scieGap"] , tmp[,"GGI"] , method="spearman"))
    
    ## return correlation coefficient
    return(cor$estimate)

}

# run a for loop to get all 10 estimates
df <- c()
for (i in 1:10) {
  
  df[i] <- get.cor1(dat.s, i)
  
}

# average the estimates
mean(df) # 0.4045611
```

Second, we try a different way of calculating relative strength. We first calculate the relative strength by subtracting science score from the mean score of science, reading and math for each individual. We then standardize this score within each country.

```{r analysis 2}

get.cor2 <- function(dat.s, i){

    dat.s$scie <- dat.s[,paste("PV",i,"SCIE",sep="")]
    dat.s$math <- dat.s[,paste("PV",i,"MATH",sep="")]
    dat.s$read <- dat.s[,paste("PV",i,"READ",sep="")]
    
    dat.s$general <- rowMeans( dat.s[,c("math","read","scie")] ) 
    
    dat.s$iiScie <- dat.s$scie - dat.s$general
    dat.s$iiMath <- dat.s$math - dat.s$general
    dat.s$iiRead <- dat.s$read - dat.s$general
    
    dat.s <- dat.s %>% group_by(country) %>% mutate(
      iiSciez = stdz(iiScie, weight),
      iiMathz = stdz(iiMath, weight),
      iiReadz = stdz(iiRead, weight)
    )
    
    dat.s <- select(dat.s, country, weight, gender, scie, math, read, efficacy, iiSciez, iiMathz, iiReadz)
    
    cdat.s <- dat.s %>% group_by(country) %>% dplyr::summarize(
         scieGap = weighted.mean(iiSciez[gender == "Male"], w = weight[gender == "Male"]) - weighted.mean(iiSciez[gender == "Female"], w = weight[gender == "Female"]),
         mathGap = weighted.mean(iiMathz[gender == "Male"], w = weight[gender == "Male"]) - weighted.mean(iiMathz[gender == "Female"], w = weight[gender == "Female"]),
         readGap = weighted.mean(iiReadz[gender == "Male"], w = weight[gender == "Male"]) - weighted.mean(iiReadz[gender == "Female"], w = weight[gender == "Female"])
    )
    
    ## now we load in the GGI data and combine with gap pisa results
    cdat.n <- left_join(cdat.s, cdat , by = "country")
    
    ## make data frame for easier use with cor.test
    tmp = as.data.frame( cdat.n )
    tmp = tmp[ !is.na(tmp[,"GGI"]),]
    
    ## print the correlation of Figure 3A and page 585
    cor <- suppressWarnings(cor.test( tmp[,"scieGap"] , tmp[,"GGI"] , method="spearman"))
    
    ## return correlation coefficient
    return(cor$estimate)

}

# run a for loop to get all 10 estimates
df <- c()
for (i in 1:10) {
  
  df[i] <- get.cor2(dat.s, i)
  
}

# average the estimates
mean(df) # 0.4014356
```


Third, we build a multilevel model to predict raw individual relative strength score using country-level GGGI, controlling for other individual level confounders.

```{r analysis 3}
dat.s <- dat %>% select(CNTRYID, CNTSCHID, CNTSTUID,
                            ST001D01T,ST004D01T, SMINS, 
                            TMINS, ESCS, REPEAT,
                            IMMIG, JOYSCIE, INTBRSCI, 
                            INSTSCIE,SCIEEFF,EPIST, 
                            SCIEACT, MOTIVAT, W_FSTUWT, PV1SCIE) 

names(dat.s) <- c("country", "schid", "id",
                "grade","gender", "science_time",
                "total_time","ses", "graderepeat", 
                "immigration", "sci_enjoy","sci_interest", 
                "sci_instrument","sci_efficacy","sci_epist",
                "sci_activity", "sci_motivation", "weight", "pv1"
                )

dat.s <- left_join(dat.s, cdat, by = "country")

fit1 <- lmer(pv1 ~ 1 + gender + 
              (1 | schid:country) + # level 2 random effect
              (1 + gender | country ), # level 3 random effect
           weights = weight, data = dat.s)

fit2 <- lmer(pv1 ~ 1 + gender * GGI + 
              (1 | schid:country) + # level 2 random effect
              (1 + gender | country ), # level 3 random effect
           weights = weight, data = dat.s)

fit3 <- lmer(pv1 ~ 1   + 
             science_time + total_time + immigration + grade + graderepeat + ses + gender + 
              (1 | schid:country) + # level 2 random effect
              (1 + gender | country ), # level 3 random effect
           weights = weight, data = dat.s)

fit4 <- lmer(pv1 ~ 1   + 
             science_time + total_time + immigration + grade + graderepeat + ses + gender * GGI + 
              (1 | schid:country) + # level 2 random effect
              (1 + gender | country ), # level 3 random effect
           weights = weight, data = dat.s)

screenreg(list(fit1, fit2, fit3, fit4), digits = 3)



```
```{r VZ_analysis 1}
#Would GGI predict Gendergap in ScieGap/MathGap/ReadGap?
reg0<-with(tmp,lm(scieGap~GGI))
reg0.1<-with(tmp,lm(mathGap~GGI))
reg0.2<-with(tmp,lm(readGap~GGI))
with(tmp,cov(scieGap,GGI)) #scieGap 0.005014771
with(tmp,cov(mathGap,GGI)) #mathGap -0.0006225173
with(tmp,cov(readGap,GGI)) #readGap -0.00247364

summary(reg0)
summary(reg0.1)
summary(reg0.2)
```

```{r additional analysis 2- do not use}
#Try to use standardized scores zMath, zRead,z_scie (not the Intraindividual strength), to see how the strength alone may be related to efficacy.
#No interesting findings.
# Do not run!
# Just leave here fyi


#dat.s <- dat %>% select(CNTRYID,W_FSTUWT,ST004D01T, SCIEEFF,PV1MATH:PV10SCIE) 
#names(dat.s)[c(1:4)] <- c("country","weight","gender", "efficacy")

# turn factor into numeric
#dat.s$efficacy <- as.numeric(as.character(dat.s$efficacy))

#dat.s <- dat.s[!dat.s$country %in% c( 
 # "Chinese Taipei","Puerto Rico",
 # "Dominican Republic", "Kosovo",
 # "USA (Massachusetts)", "USA (North Carolina)",
 # "Argentina (Ciudad Autónoma de Buenos)",
 # "Spain (Regions)","Spain (Regions)_duplicated_971") ,  ]
#dat.s$scie <- rowMeans(dat.s[,paste("PV",1:5,"SCIE",sep="")])
#dat.s$math <- rowMeans(dat.s[,paste("PV",1:5,"MATH",sep="")])
#dat.s$read <- rowMeans(dat.s[,paste("PV",1:5,"READ",sep="")])

#dat.s <- dat.s %>% group_by(country) %>% mutate(
 # z_scie = scale(scie),
 # z_math = scale(math),
  #z_read = scale(read)
#)

#dat.s$general = rowMeans( dat.s[,c("z_math","z_read","z_scie")] ) 

#dat.s <- dat.s %>% group_by(country) %>% mutate( gz = stdz(general, weight) )

cat(" Then, we calculated each student’s intraindivid- ual strengths by subtracting zGeneral as follows: relative science strength = zScience – zGeneral, relative math strength = zMath – zGeneral, rela- tive reading strength = zReading – zGeneral.")

#dat.s$iiScie <- dat.s$z_scie - dat.s$gz
#dat.s$iiMath <- dat.s$z_math - dat.s$gz
#dat.s$iiRead <- dat.s$z_read - dat.s$gz

#dat.s <- dat.s %>% group_by(country) %>% mutate(
  #iiSciez = scale(iiScie),
  #iiMathz = scale(iiMath),
  #iiReadz = scale(iiRead)
#)

#Cov
#with(dat.s,cov(z_scie,z_math))
#with(dat.s,cov(z_scie,z_read))
#with(dat.s,cov(z_math,z_read))

#with(dat.s,cov(iiScie,iiMath))
#with(dat.s,cov(iiScie,iiRead))
#with(dat.s,cov(iiMath,iiRead))


#testset <- select(dat.s, country, weight, gender, scie, math, read, efficacy, z_scie, z_math, z_read ,iiScie, iiMath,iiRead,gz )

#testset.s <- testset %>% group_by(country) %>% dplyr::summarize(
     #scieGap = weighted.mean(z_scie[gender == "Male"], w = weight[gender == "Male"]) - weighted.mean(z_scie[gender == "Female"], w = weight[gender == "Female"]),
     #mathGap = weighted.mean(z_math[gender == "Male"], w = weight[gender == "Male"]) - weighted.mean(z_math[gender == "Female"], w = weight[gender == "Female"]),
    # readGap = weighted.mean(z_read[gender == "Male"], w = weight[gender == "Male"]) - weighted.mean(z_read[gender == "Female"], w = weight[gender == "Female"])
#)

## now we load in the GGI data and combine with gap pisa results
#testset.n <- left_join(testset.s, cdat , by = "country")

## make data frame for easier use with cor.test
#tmp1 = as.data.frame( testset.n )
#tmp1 = tmp1[ !is.na(tmp1[,"GGI"]),]

## report the N so you can compare with the paper (this should be 62)
#cat(paste("Countries with both pisa data and gender equality data:",sum(!is.na(tmp1[,"scieGap"])&!is.na(tmp1[,"GGI"]))))

## print the correlation of Figure 3A and page 585
#print ( cor.test( tmp1[,"scieGap"] , tmp1[,"GGI"] , method="spearman") )

## print the reading correlation, as reported on page 585
#print ( cor.test( tmp1[,"readGap"] , tmp1[,"GGI"] , method="spearman") )


###############################

#reg1<-with(tmp,lm(scieGap~GGI))


#summary(reg0)
#summary(reg1)



```

```{r additional analysis 3}
#Select subset
dat.s <- dat %>% select(CNTRYID,W_FSTUWT,ST004D01T, SCIEEFF,PV1MATH:PV10SCIE) 
names(dat.s)[c(1:4)] <- c("country","weight","gender", "efficacy")

# turn factor into numeric
dat.s$efficacy <- as.numeric(as.character(dat.s$efficacy))

dat.s <- dat.s[!dat.s$country %in% c( 
  "Chinese Taipei","Puerto Rico",
  "Dominican Republic", "Kosovo",
  "USA (Massachusetts)", "USA (North Carolina)",
  "Argentina (Ciudad Autónoma de Buenos)",
  "Spain (Regions)","Spain (Regions)_duplicated_971") ,  ]
dat.s$scie <- rowMeans(dat.s[,paste("PV",1:5,"SCIE",sep="")])
dat.s$math <- rowMeans(dat.s[,paste("PV",1:5,"MATH",sep="")])
dat.s$read <- rowMeans(dat.s[,paste("PV",1:5,"READ",sep="")])

dat.s <- dat.s %>% group_by(country) %>% mutate(
  z_scie = scale(scie),
  z_math = scale(math),
  z_read = scale(read)
)

dat.s$general = rowMeans( dat.s[,c("z_math","z_read","z_scie")] ) 

dat.s <- dat.s %>% group_by(country) %>% mutate( gz = stdz(general, weight) )

cat("Step 3: Then, we calculated each student’s intraindivid- ual strengths by subtracting zGeneral as follows: relative science strength = zScience – zGeneral, relative math strength = zMath – zGeneral, rela- tive reading strength = zReading – zGeneral.")

dat.s$iiScie <- dat.s$z_scie - dat.s$gz
dat.s$iiMath <- dat.s$z_math - dat.s$gz
dat.s$iiRead <- dat.s$z_read - dat.s$gz

dat.s <- dat.s %>% group_by(country) %>% mutate(
  iiSciez = scale(iiScie),
  iiMathz = scale(iiMath),
  iiReadz = scale(iiRead)
)

testset <- select(dat.s, country, weight, gender, scie, math, read, efficacy, z_scie, z_math, z_read ,iiScie, iiMath,iiRead,gz )

## now we load in the GGI data and combine with gap pisa results
testset.n <- left_join(testset, cdat , by = "country")

## make data frame for easier use with cor.test
tmp1 = as.data.frame( testset.n )
tmp1 = tmp1[ !is.na(tmp1[,"GGI"]),]

###############################
###############################
###############################
########analysis###############
###############################
###############################
###############################


#See if standardized strength or intraindividual strengths would affect GGI
reg2.1<-with(tmp1,lm(GGI~iiScie+iiMath+iiRead))
reg2.2<-with(tmp1,lm(GGI~iiScie*iiMath*iiRead))
reg2.3<-with(tmp1,lm(GGI~z_scie+z_math+z_read))
reg2.4<-with(tmp1,lm(GGI~z_scie*z_math*z_read))

#Summary
summary(reg2.1)
summary(reg2.2)
summary(reg2.3)
summary(reg2.4)
anova(reg2.1,reg2.2)
anova(reg2.3,reg2.4)

#Best fit model
models<-list(reg2.1,reg2.2,reg2.3,reg2.4)
ps = unlist(lapply(models, function(x) {nrow(summary(x)$coef)-1})) 
rsquares = unlist(lapply(models, function(x){summary(x)$r.squared}))
adj.rsquares = unlist(lapply(models, function(x){summary(x)$adj.r.squared})) 
loglikes = unlist(lapply(models, logLik))
AICs = unlist(lapply(models, AIC))
BICs = unlist(lapply(models, BIC))
which.max(rsquares) 
which.max(adj.rsquares)
which.max(loglikes)
which.min(AICs)
which.min(BICs)

#Check q-q plot
plot(reg2.2)

```

